---
title: "Assignment 3"
author: "Matthew McAvoy"
date: "March 6, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and look for anomalies

```{r, include=FALSE}
library(tidyverse)
library(stringr)

setwd("C:/Users/homur/OneDrive/New College/Visualization/Week4")
```

```{r, include=FALSE}
data <- read_csv("fl_uni_employee_salary.csv")

names(data) <- c("University", "Budget.Entity", "Pos.Num", "Last.N", "First.N", "MI", "Employee.Type", "FTE", "Class.Code", "Class.Title", "Annual.Salary", "OPS.Term")
```

Lets first take a look at the data
```{r}
summary(data)
```

There are a number of NA's, especially in OPS Term Amount and Annual Salary. This makes sense as a person can only be one or the other.

Next is to check if there are duplicate people in the data set.

```{r}
repeats <- duplicated(data)
uniqData <- data[!repeats,]

sum(repeats); dim(data); dim(uniqData)
```

We see there are 7250 duplicated entries, or about 8% of the data.

## Salary data NCF vs SUS

Lets find New College's code

```{r}
unique(uniqData$University)
```

We will want to group by NCF and put everyone else in SUS. We can make a new column with only those categories then group and summarize on salary.

```{r}
ud <- uniqData %>% mutate(NCFSUS = ifelse(University=="NCF", "NCF", "SUS"))

ud2 <- ud %>% filter(!is.na(Annual.Salary)) %>% group_by(NCFSUS) %>%
  summarise(avgSalary=mean(Annual.Salary))
```

Bar chart looking at Average salary for all employees at NCF and SUS at large.

```{r}
ggplot(data=ud2, aes(x=NCFSUS, y=avgSalary)) + geom_bar(aes(fill=NCFSUS), stat="identity") + 
  theme_minimal()  + labs(title="Aggregate Salaries between NCF and SUS Employees", x="Group", y="Annual Salary") + 
  scale_fill_discrete(name="Groups", labels=c("NCF", "SUS"))
```

Looks like NCF is paying their employees less.

## NCF vs SUS by Class Title

Lets separate by class title now. There are quite alot of titles. Lets group them based on the top nine. The tenth is NA and after that not as important.

```{r}
topTitles <- ud %>% group_by(Class.Title) %>% 
  summarise(mostTitle = n()) %>% select(Class.Title, mostTitle) %>%
  arrange(desc(mostTitle)) %>% head(9)
```

We will only want to look at annual salaried employees that are one of the class types we want to investigate.

```{r}
ud3 <- ud %>% filter(!is.na(Annual.Salary)) %>% filter(Class.Title %in% topTitles$Class.Title) 

unique(ud3$Class.Title)
```

This reduces the class titles to six. A resident is not something I'm particularly familiar with, so I wll drop anyone that is a resident. Lecturer and Instructor to me seem to be the same, so I will label them both as Lecturer.

```{r}
ud3 <- ud3 %>% filter(Class.Title != "RESIDENT")

ud3$Class.Title <- sub("INSTRUCTOR", "LECTURER", ud3$Class.Title)

ud3 %>% distinct(Class.Title)
```

We are now ready to look at some graphics.

## Annual Salary among all Florida Professors

```{r}
ggplot(data=ud3, aes(x=Class.Title, y=Annual.Salary)) +
  geom_boxplot(aes(color=Class.Title)) + theme_minimal() + theme(axis.text.x = element_blank()) + 
  labs(title="Salary Among Florida Professors", x="Class Title", y="Annual Salary") + 
  scale_color_discrete(name="Class Title", labels=c("LECTURER", "ASSISTANT PROFESSOR", "ASSOCIATE PROFESSOR", "PROFESSOR"))
```

It Looks like Professors make the most out of teaching faculty.

## Comparison of Salary among NCF Professors and SUS Professors.

Changed facet grid names to fit on page and factor class titles to put in order I want.

```{r}
Title_names <- c(
                    `ASSISTANT PROFESSOR` = "PROF1",
                    `ASSOCIATE PROFESSOR` = "PROF2",
                    `PROFESSOR` = "PROF3",
                    `LECTURER` = "LECT"
                    )

ud3$Class.Title <- factor(ud3$Class.Title, levels = c("LECTURER", "ASSISTANT PROFESSOR", "ASSOCIATE PROFESSOR", "PROFESSOR"))
```


```{r}
ggplot(data=ud3, aes(x=NCFSUS, y=Annual.Salary)) + geom_boxplot(aes(color=NCFSUS)) + 
  facet_grid(~Class.Title, labeller = as_labeller(Title_names)) + theme_minimal() + 
  theme(axis.text.x = element_blank()) + 
  labs(title="Salaries of NCF and SUS Professors", x="Professor level", y="Annual Salary") + 
  scale_color_discrete(name="Groups", labels=c("NCF", "SUS"))
```

With such a large tail, lets filter on a top salary of 300,000, as well as exclude anyone making less than 10,000. These might be coding errors.

```{r}
ud4 <- ud3 %>% filter(Annual.Salary < 300000 & Annual.Salary > 10000)
```

```{r}
ggplot(data=ud4, aes(x=NCFSUS, y=Annual.Salary)) + geom_boxplot(aes(color=NCFSUS)) + 
  facet_grid(~Class.Title, labeller = as_labeller(Title_names)) + theme_minimal() + 
  theme(axis.text.x = element_blank()) + 
  labs(title="Salaries of NCF and SUS Professors", x="Professor level", y="Annual Salary") + 
  scale_color_discrete(name="Groups", labels=c("NCF", "SUS"))
```

It looks like NCF pays their professors less than most schools in the rest of Florida.

## Pay difference between administration and professors

We want to next look at the difference in pay between administrative and faculty salaries. Looking at possible titles, I will chose the following for labelling as faculty or administration

Faculty
 * Professor
 * Assistant Professor
 * Associate Professor

Administration
 * Faculty Administrator
 * Office Manager
 * Assistant Director
 * Director

I will begin by aggregating to take a big picture look at how the two vary, compare them separately, then compare Professors and Directors at NCF.

## Faculty vs Administration pay

```{r}
FacMin <- c("PROFESSOR", "ASSISTANT PROFESSOR", "ASSOCIATE PROFESSOR", "FACULTY ADMINISTRATOR", "OFFICE MANAGER", "ASSISTANT DIRECTOR", "DIRECTOR")

FAD <- ud %>% filter(!is.na(Annual.Salary)) %>% filter(Class.Title %in% FacMin) %>% 
  filter(Annual.Salary < 300000 & Annual.Salary > 10000)
```

Adding Group titles

```{r}
FacTitles <- c("PROFESSOR", "ASSISTANT PROFESSOR", "ASSOCIATE PROFESSOR")

FAD <- FAD %>% mutate(Group = ifelse(Class.Title %in% FacTitles, "FACULTY", "ADMINISTRATION"))
```

```{r}
ggplot(data=FAD, aes(x=Group, y=Annual.Salary)) + 
  geom_boxplot(aes(color=Group)) + theme_minimal() + theme(axis.text.x = element_blank()) + 
  labs(title="Salaries of Faculty vs Administration", x="Groups", y="Annual Salary") + 
  scale_color_discrete(name="Groups", labels=c("Administration", "Faculty"))
```

It looks like faculty may be payed more than administration.

Lets look at how it splits for NCF.

```{r}
FAD_NCF <- FAD %>% filter(University == "NCF")
```

```{r}
ggplot(data=FAD_NCF, aes(x=Group, y=Annual.Salary)) + 
  geom_boxplot(aes(color=Group)) + theme_minimal() + theme(axis.text.x = element_blank()) + 
  labs(title="Salaries of Faculty vs Administration at NCF", x="Groups", y="Annual Salary") + 
  scale_color_discrete(name="Groups", labels=c("Administration", "Faculty"))
```

The difference is more pronounced with faculty having higher salaries than administration.

## Classes Salary Pay

Next is to look at how pay varies by different class titles.

```{r}
FAD$Class.Title <- factor(FAD$Class.Title, levels = c("OFFICE MANAGER", "FACULTY ADMINISTRATOR", "ASSISTANT DIRECTOR", "DIRECTOR", "ASSISTANT PROFESSOR", "ASSOCIATE PROFESSOR", "PROFESSOR"))

ggplot(data=FAD, aes(x=Class.Title, y=Annual.Salary)) + 
  geom_boxplot(aes(color=Class.Title)) + theme_minimal() + theme(axis.text.x = element_blank()) + 
  labs(title="Salaries of Faculty and Administration", x="Class Levels", y="Annual Salary") + 
  scale_color_discrete(name="Class Levels")
```

Faculty administrators seem to be payed pretty well. Now we can look at NCF specifically.

## Classes Salary Pay at NCF

```{r}
FAD_NCF$Class.Title <- factor(FAD_NCF$Class.Title, levels = c("OFFICE MANAGER", "FACULTY ADMINISTRATOR", "ASSISTANT DIRECTOR", "DIRECTOR", "ASSISTANT PROFESSOR", "ASSOCIATE PROFESSOR", "PROFESSOR"))

ggplot(data=FAD_NCF, aes(x=Class.Title, y=Annual.Salary)) + 
  geom_boxplot(aes(color=Class.Title)) + theme_minimal() + theme(axis.text.x = element_blank()) + 
  labs(title="Salaries of Faculty and Administration at NCF", x="Class Levels", y="Annual Salary") + 
  scale_color_discrete(name="Class Levels")
```

Overall, it looks like Professors all over Florida make more than most administrators. 

Caution should be noted in looking at SUS data; some of the values seem abmornally high. I can't rule them out as being too high, but is sceptical.






